Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-01-26 paper/paper.tex
second version submitted

\maketitle
\begin{abstract}
Single transferrable vote (STV) is a family of preferential voting systems, different instances of which are used in binding elections throughout the world. We give a formal specification of this family, from which we derive fully verified tools that verify the computation for various instances of STV vote counting. These tools validate the probably correct execution of a run of a vote counting algorithm, based on a transcript of the count.

Our framework distils the similarities and differences of various instances of STV and gives a uniform and modular way of synthesising verifiers for its various instances, and provides the flexibility and ease for adapting and extending it to a variety of STV schemes. We minimise the trusted base in correctness of the tools produced by using the HOL4 and CakeML as the technical basis. We first formally specify and verify the tools in HOL4 and then obtain the machine executable versions for the tools by relying on the verified proof translator and the compiler of the CakeML. Moreover, proofs that we establish in HOL4 and CakeML are almost completely automated so that new verified instances of STV can be created with no (or minimal) extra proof. Finally, our experimental results with executable code demonstrate feasibility of deploying the framework for verifying real size elections having an STV counting algorithm.
\end{abstract}

@@

We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and then verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack enough skills in grasping functional programming style which we rely on for implementing verifiers in HOL4. However, the same user most probably has had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, facilitates the users with means to understand the functionality of components, modules and the system as a whole simply by inspecting the specifications instead of implementations.
\subsection{Data Structure of Machine States}
\label{MachineData}
 We choose the data structure  given in Definition~\ref{judgement} for implementing the abstract data type underlying the generic STV.  We have four reasons for this choice of data structure. 
 \begin{itemize}
\item As the abstract syntactic representation of evidence closely represents concrete evidences, designing the parser and inspecting  its  correctness is less challenging.  
 \item HOL4 has well developed libraries comprised of verified assertions of operations on list structure. By structuring the data on lists, we facilitate us and users  with exploitation of  already verified tools to formalise the framework and avoid inventing the unnecessary from the scratch. 
 \item HOL4 also has well-developed tactics for discharging proof obligations on assertions involving list structure and operations on it. We therefore come to provide us with means for ease of verification of the formalised assertions.
 \item Understanding the data structure used in the  framework implementation is critical for ease of  its usability and extensibility by third parties.  The type $\mathit{judgement}$ closely models concrete evidences such as the one in figure~\ref{EvInst}. Therefore one can sensibly perceive the  abstraction step taken for modelling  concrete data which in turn enhances understandability of the framework and its mechanism. 
   \end{itemize}
\begin{definition}\label{judgement}
We formalise machine states as an inductive type  $\mathit{judgement}$ whose constructors are  $\mathit{NonFinal}$ and $\mathit{Final}$; 
\begin{holthmenv}
\HOLTyOp{judgement}\;=\\
\;\;\;\;\HOLConst{NonFinal}\;(\HOLTyOp{ballots}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{tallies}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{piles}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list})\\
\;\;\HOLTokenBar{}\;\HOLConst{Final}\;(\HOLTyOp{cand}\;\HOLTyOp{list})
\end{holthmenv}

 A $\mathit{Final~(w)}$ judgement declares $w$ as winners of the election. A  $\mathit{NonFinal~(ba,t,p,bl_{1},bl_{2},e,h)}$ judgement  consists of uncounted ballots $ba$, tally $t$, pile $p$, $bl_{1}$ the backlog of elected candidates, $bl_{2}$ the backlog of eliminated candidates, and $e$ and $h$ for the list of elected and continuing~(hopeful) candidates. The types  $\mathit{ballots}$, $\mathit{tallies}$ and $\mathit{piles}$ are respectively abbreviations for  $\mathit{((cand ~list)\times rat) list}$,   $\mathit{(cand\times rat)~list}$ and $\mathit{(cand\times(ballots~list)~list) ~list}$ where $\mathit{rat}$ is the HOL4 type of fractional numbers.
\end{definition}
 Recall that objects whose type is $\mathit{piles}$ serve as containers for recording the ballots allocated to each candidate. Our choice for type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $\mathit{(cand\times(ballots~list))~list}$;  
\paragraph*{(a)} Some STV schemes such as the lower house ACT and Tasmania STV, employ a nation called ''last parcel''. In short the last parcel of a candidate is the collection of those ballots received by the candidate at the last round of application of the count transition which made the tally of the candidate reach or exceed the quota and therefore be elected.  Only the last parcel, instead of all, of an elected candidate is then distributed the ballots received by the candidate. Also  they update the fractional transfer value, that each ballot in the last parcel  awaiting transfer carries, based on the length of the last parcel. To  accommodate this notion and its effect, we need identifying the ballots constituting the last parcel of an elected. We hence formalise the pile to record, for each  continuing candidate, a list consisting of lists of ballots each received upon  chronological applications of count transition.  Once a candidate is elected their pile may be manipulated differently from the continuing candidates. 
\paragraph*{(b)} We noted earlier that the STV used for the upper house elections in the Victoria state transfers votes of an eliminated candidate chunk by chunk in several applications of transfer-removed, rather than in a single action, in order of the magnitude of the fractional value that the chunks carry. We therefore need to rearrange the pile of ballots of an eliminated candidate into lists based on equality of the fractional value of ballots and then distribute them in the proper order as the scheme requires. This means that the type of piles has to be as it is defined above.   
Implementing piles in this way enables us to tailor the semantics of the transfer and elect transitions and their instantiations in such a way to modularly formalise several STV schemes which use the last parcel or stepwise distribution of votes form eliminated candidates or the combination of both.

@@

Similarly  
we obtain specification and computational implementation for other machine transitions as well. Then the specification (implementation) of the machine semantics is comprised of the collection of specification (resp. implementation) of the individual transitions. We formally prove 
%formally prove that for 
each transition implementation  matches with its specification.
\begin{theorem}
Assume $M^{spec} = \langle \mathcal{S}, \mathcal{T}, (S_{t}^{spec})_{t \in \mathcal{T}} \rangle$ is the specification of the machine and  $M^{imp} = \langle \mathcal{S}, \mathcal{T}, (S_{t}^{dec})_{t \in \mathcal{T}} \rangle$ is its computational implementation. Then for any $t\in\mathcal{T}$,  $\mathcal{S}_{t}^{spec}\Leftrightarrow\mathcal{S}_{t}^{dec}$.      
\end{theorem}

@@

\footnote{The Tasmania STV also uses similar surplus transfer mechanism.}. Instantiation of other machine transitions proceed in a  similar manner.    

@@

\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l\sp{\prime}}\;\HOLBoundVar{c}.\\
\;\;\;\;\;\;\;\;\;\;(\HOLBoundVar{bl}\;\HOLSymConst{=}\;\HOLBoundVar{c}\HOLSymConst{::}\HOLBoundVar{l\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\

@@

\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d\sp{\prime}}.\\
\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLBoundVar{d\sp{\prime}}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l\sp{\prime}}.\\
\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;(\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{np})\;\HOLSymConst{\HOLTokenConj{}}\\

@@

 \HOLConst{less_than_quota}\;\HOLFreeVar{qu}\;\HOLFreeVar{l}\;\HOLFreeVar{ls}\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{EVERY}\;(\HOLTokenLambda{}\HOLBoundVar{h}.\;\HOLConst{get_cand_tally}\;\HOLBoundVar{h}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{qu})\;\HOLFreeVar{ls}

@@

\;\;\;\;(\HOLKeyword{let}\\
\;\;\;\;\;\;\;\HOLBoundVar{gcp}\;=\;\HOLConst{get_cand_pile}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p}\\
\;\;\;\;\;\HOLKeyword{in}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{hbl}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl\sp{\prime}}\;\HOLSymConst{=}\;\HOLBoundVar{tbl}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{last}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\\

@@

\begin{itemize}
\item[$\bullet_{6'}$] transfer all of the surplus votes of an elected candidate at a reduced fraction.
\item[$\bullet_{8'}$] the fractional transfer value is computed based on all of the surplus (not necessarily depending on the last parcel).
\end{itemize}

Note that we deal with updating fractional transfer value in the semantics of elect transitions. Also in instantiations of the elect semantics, we guarantee elements of the backlog of elected candidates ($bl_{1}$) are all above the quota. Therefore we formalise Victoria's transfer-elect semantics as follows.
  \HOLConst{TransferVicDec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;(\HOLConst{NonFinal}\;(\HOLFreeVar{ba}\HOLSymConst{,}\HOLFreeVar{t}\HOLSymConst{,}\HOLFreeVar{p}\HOLSymConst{,}\HOLFreeVar{bl}\HOLSymConst{,}\HOLFreeVar{bl\sb{\mathrm{2}}}\HOLSymConst{,}\HOLFreeVar{e}\HOLSymConst{,}\HOLFreeVar{h}))\\
\;\;\;\;(\HOLKeyword{let}\\
\;\;\;\;\;\;\;\HOLBoundVar{gcp}\;=\;\HOLConst{get_cand_pile}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p}\\
\;\;\;\;\;\HOLKeyword{in}\\
\;\;\;\;\;\;\;\HOLFreeVar{bl\sp{\prime}}\;\HOLSymConst{=}\;\HOLBoundVar{tbl}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{flat}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{hbl}\HOLSymConst{,}\HOLConst{\ensuremath{[\,]}})\;\HOLFreeVar{p\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\HOLConst{subpile1}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p}\;\HOLFreeVar{p\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{subpile2}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{p})\\
\HOLConst{TransferVicDec}\;\HOLFreeVar{v\sb{\mathrm{0}}}\;(\HOLConst{Final}\;\HOLFreeVar{v\sb{\mathrm{1}}})\;\HOLFreeVar{v\sb{\mathrm{2}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{false}\\
\HOLConst{TransferVicDec}\;\HOLFreeVar{v\sb{\mathrm{3}}}\;(\HOLConst{NonFinal}\;\HOLFreeVar{v\sb{\mathrm{9}}})\;(\HOLConst{Final}\;\HOLFreeVar{v\sb{\mathrm{5}}})\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{false}

@@

\;\;\;\;\HOLBoundVar{tbl}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{APPEND_ALL}\;\HOLFreeVar{p}\;\HOLSymConst{\HOLTokenConj{}}\\

@@

\;\;\HOLConst{Valid_Step}\;\HOLFreeVar{params}\;\HOLFreeVar{j\sb{\mathrm{0}}}\;\HOLFreeVar{j\sb{\mathrm{1}}}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{valid_judgements_dec}\;\HOLFreeVar{params}\;(\HOLFreeVar{j\sb{\mathrm{1}}}\HOLSymConst{::}\HOLFreeVar{js})

@@

\HOLConst{Check_Parsed_Certificate}\;\HOLFreeVar{params}\\
\;\;(\HOLFreeVar{first\HOLTokenUnderscore{}judgement}\HOLSymConst{::}\HOLFreeVar{rest\HOLTokenUnderscore{}judgements})\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{Initial_Judgement_dec}\;(\HOLConst{snd}\;(\HOLConst{snd}\;\HOLFreeVar{params}))\;\HOLFreeVar{first\HOLTokenUnderscore{}judgement}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{valid_judgements_dec}\;\HOLFreeVar{params}\;(\HOLFreeVar{first\HOLTokenUnderscore{}judgement}\HOLSymConst{::}\HOLFreeVar{rest\HOLTokenUnderscore{}judgements})

>>>>

\maketitle
\begin{abstract}
Single transferrable vote (STV) is a family of preferential voting systems, different instances of which are used in binding elections throughout the world. We give a formal specification of this family, from which we derive fully verified tools that verify the computation for various instances of STV vote counting. These tools validate the probably correct execution of a run of a vote counting algorithm, based on a transcript of the count.
 Our framework distils the similarities and differences of various instances of STV and gives a uniform and modular way of synthesising verifiers for its various instances, and provides the flexibility and ease for adapting and extending it to a variety of STV schemes. We minimise the trusted base in correctness of the tools produced by using the HOL4 and CakeML as the technical basis. We first formally specify and verify the tools in HOL4 and then obtain the machine executable versions for the tools by relying on the verified proof translator and the compiler of the CakeML. Moreover, proofs that we establish in HOL4 and CakeML are almost completely automated so that new verified instances of STV can be created with no (or minimal) extra proof. Finally, our experimental results with executable code demonstrate feasibility of deploying the framework for verifying real size elections having an STV counting algorithm.
\end{abstract}

@@

We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and then verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack enough skills in grasping functional programming style which we rely on for implementing verifiers in HOL4. However, the same user most probably has had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, facilitates the users with means to understand the functionality of components, modules and the system as a whole simply by inspecting the specifications instead of implementations.
\subsection{Data Structure of Machine States}
\label{MachineData}
 We choose the data structure  given in Definition~\ref{judgement} for implementing the abstract data type underlying the generic STV.  We have four reasons for this choice of data structure. 
 \begin{itemize}
\item As the abstract syntactic representation of evidence closely represents concrete evidences, designing the parser and inspecting  its  correctness is less challenging.  
 \item HOL4 has well developed libraries comprised of verified assertions of operations on list structure. By structuring the data on lists, we facilitate us and users  with exploitation of  already verified tools to formalise the framework and avoid inventing the unnecessary from the scratch. 
 \item HOL4 also has well-developed tactics for discharging proof obligations on assertions involving list structure and operations on it. We therefore come to provide us with means for ease of verification of the formalised assertions.
 \item Understanding the data structure used in the  framework implementation is critical for ease of  its usability and extensibility by third parties.  The type $\mathit{judgement}$ closely models concrete evidence such as figure~\ref{EvInst}. Therefore one can sensibly perceive the  abstraction step taken for modelling  concrete data which in turn enhances understandability of the framework and its mechanism. 
   \end{itemize}
\begin{definition}\label{judgement}
We formalise machine states as an inductive type  $\mathit{judgement}$ whose constructors are  $\mathit{NonFinal}$ and $\mathit{Final}$; 
\begin{holthmenv}
\HOLTyOp{judgement}\;=\\
\;\;\;\;\HOLConst{NonFinal}\;(\HOLTyOp{ballots}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{tallies}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{piles}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list})\\
\;\;\HOLTokenBar{}\;\HOLConst{Final}\;(\HOLTyOp{cand}\;\HOLTyOp{list})
\end{holthmenv}

 A $\mathit{Final~(w)}$ judgement declares $w$ as winners of the election. A  $\mathit{NonFinal~(ba,t,p,bl_{1},bl_{2},e,h)}$ judgement  consists of uncounted ballots $ba$, tally $t$, pile $p$, $bl_{1}$ the backlog of elected candidates, $bl_{2}$ the backlog of eliminated candidates, and $e$ and $h$ for the list of elected and continuing~(hopeful) candidates. The types  $\mathit{ballots}$, $\mathit{tallies}$ and $\mathit{piles}$ are respectively abbreviations for  $\mathit{((cand ~list)\times rat) list}$,   $\mathit{(cand\times rat)~list}$ and $\mathit{(cand\times(ballots~list)~list) ~list}$ where $\mathit{rat}$ is the HOL4 type of fractional numbers.
\end{definition}
 Recall that objects whose type is $\mathit{piles}$ serve as containers for recording the ballots allocated to each candidate. Our choice for type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $\mathit{(cand\times(ballots~list))~list}$;  
\paragraph*{(a)} Some STV schemes such as  the ACT and Tasmania STV, employ a nation called ''last parcel''. In short the last parcel of a candidate is the collection of those ballots received at the last round of application of the count transition which made the tally of the candidate reach or exceed the quota and therefore be elected.  Only the last parcel, instead of all, of an elected candidate is then distributed the ballots received by the candidate. Also  they update the fractional transfer value, that each ballot in the last parcel  awaiting transfer carries, based on the length of the last parcel. To  accommodate this notion and its effect, we need identifying the ballots constituting the last parcel of an elected. We hence formalise the pile to record, for each  continuing candidate, a list consisting of lists of ballots each received upon  chronological applications of count transition.  Once a candidate is elected their pile may be manipulated differently from the continuing candidates. 
\paragraph*{(b)} We noted earlier that the STV used for the upper house elections in the Victoria state transfers votes of an eliminated candidate chunk by chunk in several applications of transfer-removed, rather than in a single action, in order of the magnitude of the fractional value that the chunks carry. We therefore need to rearrange the pile of ballots of an eliminated candidate into lists based on equality of the fractional value of ballots and then distribute them in the proper order as the scheme requires. This means that the type of piles has to be as it is defined above. Implementing piles in this way enables
% us to 
tailoring the semantics of the transfer and elect transitions and their instantiations in such a way to modularly formalise several STV schemes which use the last parcel and/or stepwise distribution of votes from eliminated candidates. 
%or the combination of both.

@@

Similarly  
we obtain specification and computational implementation for other machine transitions as well. Then the specification (implementation) of the machine semantics is comprised of the collection of specification (resp. implementation) of the individual transitions. We formally prove 
%formally prove that for 
each transition implementation  matches with its specification.
\begin{theorem}
Assume $M^{spec} = \langle \mathcal{S}, \mathcal{T}, (S_{t}^{spec})_{t \in \mathcal{T}} \rangle$ and $M^{imp} = \langle \mathcal{S}, \mathcal{T}, (S_{t}^{dec})_{t \in \mathcal{T}} \rangle$ are the specification and  implementation of the machine. Then  $\forall t\in\mathcal{T}$,  $\mathcal{S}_{t}^{spec}\Leftrightarrow\mathcal{S}_{t}^{dec}$.      
\end{theorem}

@@

%\footnote{The Tasmania STV also uses similar surplus transfer mechanism.}. 
Instantiation of other machine transitions proceed in a  similar manner.    

@@

\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l\sp{\prime}}\;\HOLBoundVar{c}.
\;\;(\HOLBoundVar{bl}\;\HOLSymConst{=}\;\HOLBoundVar{c}\HOLSymConst{::}\HOLBoundVar{l\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{c}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\\

@@

\;\;\;\;\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{d\sp{\prime}}.
\;\HOLBoundVar{d\sp{\prime}}\;\HOLSymConst{\HOLTokenNotEqual{}}\;\HOLBoundVar{c}\;\HOLSymConst{\HOLTokenImp{}}
\;\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{l\sp{\prime}}.
\;\;(\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{p}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{d\sp{\prime}}\HOLSymConst{,}\HOLBoundVar{l\sp{\prime}})\;\HOLBoundVar{np})\;\HOLSymConst{\HOLTokenConj{}}\\

@@

 \HOLConst{less_than_quota}\;\HOLFreeVar{qu}\;\HOLFreeVar{l}\;\HOLFreeVar{ls}\\\;\HOLSymConst{\HOLTokenEquiv{}}
\;\HOLConst{EVERY}\;(\HOLTokenLambda{}\HOLBoundVar{h}.\;\HOLConst{get_cand_tally}\;\HOLBoundVar{h}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenLt{}}\;\HOLFreeVar{qu})\;\HOLFreeVar{ls}

@@

\;\;\;\;(\HOLKeyword{let}
\;\;\;\;\;\;\;\HOLBoundVar{gcp}\;=\;\HOLConst{get_cand_pile}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p}
\;\;\;\HOLKeyword{in}\\
\;\;\;\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;\HOLBoundVar{hbl}\;\HOLFreeVar{l}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl\sp{\prime}}\;\HOLSymConst{=}\;\HOLBoundVar{tbl}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{last}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\\

@@

\begin{itemize}
\item[$\bullet_{6'}$] transfer all of the surplus votes of an elected candidate at a reduced fraction.
\item[$\bullet_{8'}$] the fractional transfer value is computed based on all of the surplus (not necessarily depending on the last parcel).
\end{itemize} 
\begin{small}
 \begin{holthmenv}
  \HOLConst{TransferVicDec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;(\HOLConst{NonFinal}\;(\HOLFreeVar{ba}\HOLSymConst{,}\HOLFreeVar{t}\HOLSymConst{,}\HOLFreeVar{p}\HOLSymConst{,}\HOLFreeVar{bl}\HOLSymConst{,}\HOLFreeVar{bl\sb{\mathrm{2}}}\HOLSymConst{,}\HOLFreeVar{e}\HOLSymConst{,}\HOLFreeVar{h}))\\
\;\;(\HOLConst{NonFinal}\;(\HOLFreeVar{ba\sp{\prime}}\HOLSymConst{,}\HOLFreeVar{t\sp{\prime}}\HOLSymConst{,}\HOLFreeVar{p\sp{\prime}}\HOLSymConst{,}\HOLFreeVar{bl\sp{\prime}}\HOLSymConst{,}\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\HOLSymConst{,}\HOLFreeVar{e\sp{\prime}}\HOLSymConst{,}\HOLFreeVar{h\sp{\prime}}))\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{TransferAuxDec}\;(\HOLFreeVar{qu}\HOLSymConst{,}\HOLFreeVar{st}\HOLSymConst{,}\HOLFreeVar{l})\;\HOLFreeVar{ba}\;\HOLFreeVar{t}\;\HOLFreeVar{t\sp{\prime}}\;\HOLFreeVar{p}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{bl}\;\HOLFreeVar{bl\sb{\mathrm{2}}}\;\HOLFreeVar{bl\sb{\mathrm{2}}\sp{\prime}}\;\HOLFreeVar{e}\;\HOLFreeVar{e\sp{\prime}}\;\HOLFreeVar{h}\;\HOLFreeVar{h\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{less_than_quota}\;\HOLFreeVar{qu}\;\HOLFreeVar{t}\;\HOLFreeVar{h}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;dtcase\;\HOLFreeVar{bl}\;\HOLKeyword{of}\\
\;\;\;\;\HOLConst{\ensuremath{[\,]}}\;\HOLTokenImp{}\;\HOLConst{false}\\
\;\;\HOLTokenBar{}\;\HOLBoundVar{hbl}\HOLSymConst{::}\HOLBoundVar{tbl}\;\HOLTokenImp{}
\;\;(\HOLKeyword{let}
\;\;\HOLBoundVar{gcp}\;=\;\HOLConst{get_cand_pile}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p}
\;\;\;\HOLKeyword{in}\\
\;\;\;\;\;\;\;\HOLFreeVar{bl\sp{\prime}}\;\HOLSymConst{=}\;\HOLBoundVar{tbl}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLSymConst{\HOLTokenNeg{}}\HOLConst{null}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{flat}\;\HOLBoundVar{gcp}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{\HOLConst{mem}}\;(\HOLBoundVar{hbl}\HOLSymConst{,}\HOLConst{\ensuremath{[\,]}})\;\HOLFreeVar{p\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\HOLConst{subpile1}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p}\;\HOLFreeVar{p\sp{\prime}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{subpile2}\;\HOLBoundVar{hbl}\;\HOLFreeVar{p\sp{\prime}}\;\HOLFreeVar{p})\\
\HOLConst{TransferVicDec}\;\HOLFreeVar{v\sb{\mathrm{0}}}\;(\HOLConst{Final}\;\HOLFreeVar{v\sb{\mathrm{1}}})\;\HOLFreeVar{v\sb{\mathrm{2}}}\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{false}\\
\HOLConst{TransferVicDec}\;\HOLFreeVar{v\sb{\mathrm{3}}}\;(\HOLConst{NonFinal}\;\HOLFreeVar{v\sb{\mathrm{9}}})\;(\HOLConst{Final}\;\HOLFreeVar{v\sb{\mathrm{5}}})\;\HOLSymConst{\HOLTokenEquiv{}}\;\HOLConst{false}
\end{holthmenv}
\end{small}
Note that we deal with updating fractional transfer value in the semantics of elect transitions. Also in instantiations of the elect semantics, we guarantee elements of the backlog of elected candidates ($bl_{1}$) are all above the quota. Therefore we formalise Victoria's transfer-elect semantics as above.

@@

\;\;\HOLTokenBar{}\;\HOLBoundVar{hbl}\HOLSymConst{::}\HOLBoundVar{tbl}\;\HOLTokenImp{}
\;\;\HOLBoundVar{tbl}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{bl\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{\ensuremath{[\,]}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLFreeVar{ba\sp{\prime}}\;\HOLSymConst{=}\;\HOLConst{APPEND_ALL}\;\HOLFreeVar{p}\;\HOLSymConst{\HOLTokenConj{}}\\

@@

\;\HOLConst{Valid_Step}\;\HOLFreeVar{params}\;\HOLFreeVar{j\sb{\mathrm{0}}}\;\HOLFreeVar{j\sb{\mathrm{1}}}\;\HOLSymConst{\HOLTokenConj{}}
\;\HOLConst{valid_judgements_dec}\;\HOLFreeVar{params}\;(\HOLFreeVar{j\sb{\mathrm{1}}}\HOLSymConst{::}\HOLFreeVar{js})

@@

\HOLConst{Check_Parsed_Certificate}\;\HOLFreeVar{params}
\;\;(\HOLFreeVar{state}\HOLSymConst{::}\HOLFreeVar{rest\HOLTokenUnderscore{}states})\;\HOLSymConst{\HOLTokenEquiv{}}\\
\;\;\HOLConst{Initial_Judgement_dec}\;(\HOLConst{snd}\;(\HOLConst{snd}\;\HOLFreeVar{params}))\;\HOLFreeVar{first\HOLTokenUnderscore{}state}\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\HOLConst{valid_judgements_dec}\;\HOLFreeVar{params}\;(\HOLFreeVar{state}\HOLSymConst{::}\HOLFreeVar{rest\HOLTokenUnderscore{}states})
